---
title: "Ban400 home exam, fall 2022, Candidate no 115"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = F)
```

# Stefanski pandemic data

## Problem 1

```{r, eval=T, error=TRUE, message=TRUE}

## Insert your code for solving problem 1 here
library(tidyverse)
library(magrittr)
library(readr)
library(readxl)


defaultW <- getOption("warn")
options(warn = -1)

# Load data, select relevant variables:
df <-
  read_excel("Stefanski-pandemic-data.xlsx") %>%
  select(-tree_ring, -pandemic_pc, -gdp)

runreg <-
function(h, df) { df %>%
      select(country, year, ln_gdp, pandemic_01, war) %>%
      mutate(country = factor(country)) %>%
      group_by(country) %>%
      mutate(
        lead_ln_gdp = lead(ln_gdp, h, order_by = year),
        lag_1_pandemic_01 = lag(pandemic_01, 1, order_by = year),
        lag_1_war = lag(war, 1, order_by = year),
        lag_ln_gdp = lag(ln_gdp, 1, order_by = year),
        vienna = case_when(year >= 1815 ~ 1, TRUE ~ 0),
        ww2 = case_when(year >= 1945 ~ 1, TRUE ~ 0),
        oil = case_when(year >= 1973 ~ 1, TRUE ~ 0),
        vienna_t = vienna * year,
        ww2_t = ww2 * year,
        oil_t = oil * year
      ) %>%
      ungroup() %>%
      lm(
        lead_ln_gdp ~
          country + lag_ln_gdp + pandemic_01 +
          lag_1_pandemic_01 + year + war +
          lag_1_war + vienna + oil + year +
          ww2 + vienna_t + oil_t + ww2_t,
data = . ) %>%
      summary() %>%
      .$coefficients %>%
      as.data.frame() %>%
      rownames_to_column() %>%
      as_tibble() %>%
      mutate(h = h)
  }
```




We don't have similar result here. In S2020, we see some positive effect after 20 years. But, in the figure from the code,pandemic's effect becomes negative after approximately 24 years. I believe not considering some elements of the original calculation led to this situation.


## Problem 2

```{r, eval=T, error=TRUE, message=TRUE,warning=FALSE}

## Insert your code for solving problem 2 here
library(tidyverse)
library(readr)
library(readxl)
pandemic_data<-read_excel("Stefanski-pandemic-data.xlsx")

sim_one_country <- function(country, T, p_pand, p_war, coeff_lag_ln_gdp, sigma, impulse_response, sigma_initial_value, sigma_alpha) {
  # Initialize data frame to store simulated data
  df <- data.frame(country=character(), year=integer(), ln_gdp=numeric(), pandemic_01=numeric(), war=logical())

  # Set fixed effect (alpha)
  alpha <- rnorm(1, sd=sigma_alpha)

  # Initialize first observation of ln_gdp
  ln_gdp_t <- alpha + rnorm(1, sd=sigma_initial_value)
  M <- length(impulse_response)

  W=numeric(M)
  for(i in 1:max(T,M)){
      W[i]=rbinom(T, 1, p_pand)
  }




  # Simulate data for T years
  for (t in 1:T) {
    # Simulate pandemic and war indicators
    war <- rbinom(1, 1, p_war) == 1

    # Compute ln_gdp value
    if(t>1){
    yi_t <- alpha + coeff_lag_ln_gdp * ln_gdp_t
    
    for (h in 1:M) {
        if(t-h+1>=1){
      yi_t <- yi_t + (impulse_response[h]*W[t-h+1]) 
        }
    }
    
    yi_t <- yi_t + rnorm(1, mean=0, sd=sigma)
    }
    else{yi_t=ln_gdp_t}
    # Store simulated values in data frame
    df <- rbind(df, data.frame(country=country, year=t, ln_gdp=yi_t, pandemic_01=W[t], war=war))

    # Update ln_gdp value for next iteration
    ln_gdp_t <- yi_t
  }

  return(df)
}
simdat <- function(N, T, p_pand, p_war, sigma, coeff_lag_ln_gdp, impulse_response, sigma_alpha, sigma_initial_value) {
  # Initialize the data frame
  df <- data.frame(country=character(), year=integer(), ln_gdp=numeric(), pandemic_01=integer(), war=integer())

  # Simulate data for N countries
  for (i in 1:N) {
    # Simulate data for one country
    country_df <- sim_one_country(paste("country", i, sep=""), T, p_pand, p_war, coeff_lag_ln_gdp, sigma, impulse_response, sigma_initial_value, sigma_alpha)

    # Append the data to the data frame
    df <- rbind(df, country_df)
  }

  return(df)
}


 



```


```{r, eval=T, error=TRUE, message=FALSE,warning=FALSE}
library(assertthat)
library(magrittr)

# Tests for assignment 2. Leave this code chunk *unchanged* and 
# ensure you run the tests *after* the chunk with you answer to 
# assignment 2. 

library(assertthat)

assert_that(
  sim_one_country(
    country = 1,
    T = 5,
    p_pand = 1,
    p_war = 0,
    sigma = 0,
    coeff_lag_ln_gdp = 0,
    impulse_response = c(1),
    sigma_initial_value = 0,
    sigma_alpha = 0
  ) %>%
    filter(year > 1) %>%
    filter(ln_gdp == 1) %>%
    nrow() == 4,
  msg = "Impulse responses are not correct"
)


assert_that(
  all(
    sim_one_country(
      country = 1,
      T = 5,
      p_pand = 1,
      p_war = 0,
      sigma = 0,
      coeff_lag_ln_gdp = 0,
      impulse_response = c(1, 2, 3, 4),
      sigma_initial_value = 0,
      sigma_alpha = 0
    )$ln_gdp == c(0, 3, 6, 10, 10)
  ),
  msg = "Impulse response is not implemented correctly"
)

assert_that(
  all(
    sim_one_country(
      country = 1,
      T = 5,
      p_pand = 1,
      p_war = 0,
      sigma = 0,
      coeff_lag_ln_gdp = 0,
      impulse_response = rep(1, 10),
      sigma_initial_value = 0,
      sigma_alpha = 0
    )$ln_gdp == c(0, 2, 3, 4, 5)
  ),
  msg = "Function fails if impulse response is longer than 1:T"
)

assert_that(abs(sd(
  replicate(
    1000,
    sim_one_country(
      country = 1,
      T = 1,
      p_pand = 1,
      p_war = 0,
      sigma = 0,
      coeff_lag_ln_gdp = 0,
      impulse_response = c(0),
      sigma_initial_value = 1,
      sigma_alpha = 0
    )$ln_gdp
  )
) - 1) < .1,
msg = "Random number generation does not work"
)


assert_that(abs(mean(
  simdat(
    N = 1,
    T = 10000,
    p_pand = .1,
    p_war = 0,
    sigma = 0,
    coeff_lag_ln_gdp = 0,
    impulse_response = rep(1, 10),
    sigma_initial_value = 0,
    sigma_alpha = 0
  )$pandemic_01
) - .1) < .01,
msg = "Pandemic variable simulation does not work"
)


assert_that(
  simdat(
    N = 10,
    T = 5,
    p_pand = .1,
    p_war = 0,
    sigma = 1,
    coeff_lag_ln_gdp = 0,
    impulse_response = rep(1, 10),
    sigma_initial_value = 0,
    sigma_alpha = 0
  ) %>% nrow() == 50,
  msg = "The simdat function return wrong number of rows"
)

assert_that(
  length(unique(simdat(
    N = 100,
    T = 5,
    p_pand = .1,
    p_war = 0,
    sigma = 1,
    coeff_lag_ln_gdp = 0,
    impulse_response = rep(1, 10),
    sigma_initial_value = 0,
    sigma_alpha = 0
  )$country)) == 100,
  msg = "The simdat function return wrong number of countries"
)

```

## Problem 3

```{r, eval=T, error=TRUE, message=TRUE,warning=FALSE}

## Insert your code for solving problem 3 here
library(assertthat)
library(tidyverse)
library(magrittr)
library(readxl)


defaultW <- getOption("warn")
options(warn = -1)
library(tidyverse)
library(magrittr)
library(readr)
library(readxl)
# Load data, select relevant variables:


runreg <-
function(h, df) { df %>%
      select(country, year, ln_gdp, pandemic_01, war) %>%
      mutate(country = factor(country)) %>%
      group_by(country) %>%
      mutate(
        lead_ln_gdp = lead(ln_gdp, h, order_by = year),
        lag_1_pandemic_01 = lag(pandemic_01, 1, order_by = year),
        lag_1_war = lag(war, 1, order_by = year),
        lag_ln_gdp = lag(ln_gdp, 1, order_by = year),
        vienna = case_when(year >= 1815 ~ 1, TRUE ~ 0),
        ww2 = case_when(year >= 1945 ~ 1, TRUE ~ 0),
        oil = case_when(year >= 1973 ~ 1, TRUE ~ 0),
        vienna_t = vienna * year,
        ww2_t = ww2 * year,
        oil_t = oil * year
      ) %>%
      ungroup() %>%
      lm(
        lead_ln_gdp ~
          country + lag_ln_gdp + pandemic_01 +
          lag_1_pandemic_01 + year + war +
          lag_1_war + vienna + oil + year +
          ww2 + vienna_t + oil_t + ww2_t,
data = . ) %>%
      summary() %>%
      .$coefficients %>%
      as.data.frame() %>%
      rownames_to_column() %>%
      as_tibble() %>%
      mutate(h = h)
  }

sim_one_country <- function(country, T, p_pand, p_war, coeff_lag_ln_gdp, sigma, impulse_response, sigma_initial_value, sigma_alpha) {
  # Initialize data frame to store simulated data
  df <- data.frame(country=character(), year=integer(), ln_gdp=numeric(), pandemic_01=numeric(), war=logical())

  # Set fixed effect (alpha)
  alpha <- rnorm(1, sd=sigma_alpha)

  # Initialize first observation of ln_gdp
  ln_gdp_t <- alpha + rnorm(1, sd=sigma_initial_value)
  M <- length(impulse_response)

  W=numeric(M)
  for(i in 1:max(T,M)){
      W[i]=rbinom(T, 1, p_pand)
  }




  # Simulate data for T years
  for (t in 1:T) {
    # Simulate pandemic and war indicators
    war <- rbinom(1, 1, p_war) == 1

    # Compute ln_gdp value
    if(t>1){
    yi_t <- alpha + coeff_lag_ln_gdp * ln_gdp_t
    
    for (h in 1:M) {
        if(t-h+1>=1){
      yi_t <- yi_t + (impulse_response[h]*W[t-h+1]) 
        }
    }
    yi_t <- yi_t + rnorm(1, mean=0, sd=sigma)
    }else{yi_t=ln_gdp_t}
    # Store simulated values in data frame
    df <- rbind(df, data.frame(country=country, year=t, ln_gdp=yi_t, pandemic_01=W[t], war=war))

    # Update ln_gdp value for next iteration
    ln_gdp_t <- yi_t
  }

  return(df)
}
simdat <- function(N, T, p_pand, p_war, sigma, coeff_lag_ln_gdp, impulse_response, sigma_alpha, sigma_initial_value) {
  # Initialize the data frame
  df <- data.frame(country=character(), year=integer(), ln_gdp=numeric(), pandemic_01=integer(), war=integer())

  # Simulate data for N countries
  for (i in 1:N) {
    # Simulate data for one country
    country_df <- sim_one_country(paste("country", i, sep=""), T, p_pand, p_war, coeff_lag_ln_gdp, sigma, impulse_response, sigma_initial_value, sigma_alpha)

    # Append the data to the data frame
    df <- rbind(df, country_df)
  }

  return(df)
}





dataf_1=simdat(
N=50,
T=1000,
p_pand=.1,
p_war=0,
sigma=10,
impulse_response=c(1,-1,2,-2,0,0,1),
sigma_initial_value = 1,
sigma_alpha = 5,
coeff_lag_ln_gdp=0
)

dataf_2=simdat(
N=50,
T=1000,
p_pand=.1,
p_war=0,
sigma=10,
impulse_response=c(1,-1,2,-2,0,0,1),
sigma_initial_value = 1,
sigma_alpha = 5,
coeff_lag_ln_gdp=.96
)


# Plot the first plot
par(mfrow=c(1,2))
lapply(0:40, FUN = runreg, df = dataf_1) %>%
  bind_rows() %>%
  filter(rowname == "pandemic_01") %>%
  ggplot(aes(x = h)) +
  geom_line(aes(y = Estimate * 100)) +
  geom_ribbon(aes(ymin = (Estimate - 1.96 * `Std. Error`) * 100,
                   ymax = (Estimate + 1.96 * `Std. Error`) * 100
  ), alpha = .1) +
  geom_hline(yintercept = 0) +
  xlab("Fcast horizon") +
  ylab("Impulse response, percent") +
  ggtitle("Estimated impulse response of pandemics on GDP pr capita") +
  theme_classic()
#geom_point(data=gfg5, col="black")
# Plot the second plot
lapply(0:40, FUN = runreg, df = dataf_2) %>%
  bind_rows() %>%
  filter(rowname == "pandemic_01") %>%
  ggplot(aes(x = h)) +
  geom_line(aes(y = Estimate * 100)) +
  geom_ribbon(aes(ymin = (Estimate - 1.96 * `Std. Error`) * 100,
                   ymax = (Estimate + 1.96 * `Std. Error`) * 100
  ), alpha = .1) +
  geom_hline(yintercept = 0) +
  xlab("Fcast horizon") +
  ylab("Impulse response, percent") +
  ggtitle("Estimated impulse response of pandemics on GDP pr capita") +
  theme_classic()




```


## Problem 4

```{r, eval=T, error=TRUE, message=TRUE,warning=FALSE}

## Insert your code for solving problem 4 here



# Tests for assignment 4. Leave this code chunk *unchanged* and 
# ensure you run the tests *after* the chunk with you answer to 
# assignment 4. 


assert_that(
  runreg_baseline(0, df, 40) %>%
    filter(substr(rowname, 1, 11) == "pandemic_01") %>%
    nrow(.) == 41,
  msg = "Function does not have the right number of lags for pandemics"
)

assert_that(
  runreg_baseline(0, df, 70) %>%
    filter(substr(rowname, 1, 6) == "ln_gdp") %>%
    nrow(.) == 70,
  msg = "Function does not have the right number of lags for ln_gdp"
)
```



I tried to make a function but don't know what went wrong. Have little time to solve.


## Problem 5

```{r, eval=T, error=TRUE, message=TRUE}


library(tidyverse)
library(readxl)
library(countrycode)
library(maps)
library(gganimate)
library(ggthemes)
# Load data, select relevant variables:
df <-
  read_excel("Stefanski-pandemic-data.xlsx") %>%
 select(-tree_ring, -pandemic_01, -pandemic_pc, -ln_gdp,-war)

world <- map_data("world") %>% select(-subregion,-group,-order)



df$country <- countrycode(df$country, "iso3c", "country.name")


#here there's one code with SUN which gives NA so I'm assuming its Slovenia with SVN
df$country[is.na(df$country)] <- "Slovenia"


df_final <- left_join(df,world , by = c("country" = "region")) 

# Now we are limiting the year count because it may take too much time.
# If we want all the year in the graph, we can just comment out the line below.

#df_final <- df_final %>% filter(between(year, 1990,2016))  # This line


df_final <- df_final %>%
 group_by(year,country) %>%
 summarise(gdp = mean(gdp),long = mean(long),lat = mean(lat))



world <- ggplot() +
 borders("world", colour = "gray85", fill = "gray80") +
 theme_map() 

 world +
 geom_point(aes(x = long, y = lat, size = gdp),
            data = df_final, 
            colour = 'purple', alpha = .5) + labs(title = "Year: {frame_time}", size = "gdp") +
 transition_time(year) + ease_aes("linear")
 




```


#### Session info

Leave this part unchanged. The cell below prints which packages and versions were used for creating the html-file. 

```{r, eval=T}
sessionInfo()
```
